Fast Automatic Bayesian Cubature Using Lattice Sampling

### 2019-06-29 BayesianCubaturePaper2018 - Rev1b.tex
Text revised to reflect the randomly chosen error tolerance.

We use the following parameter values in the simulation: 
\begin{align*}
d' = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol points \cite{DicEtal14a,DicPil10a}.  The results for  $\varepsilon = 10^{-2}, 10^{-3}, 10^{-4}$, and $10^{-5}$ and $100$ scrambles for each $\varepsilon$ are shown in \figref{fig:MVN_Metern_d2b2}.  We  observe the algorithm meets the error criterion 95\% of the time.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer, 
As shown in \figref{fig:MVN_Metern_d2b2}, computation time increases rapidly with $n$. 
The  maximum likelihood estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC_\vtheta$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

@@

Three integrals were evaluated using the GAIL algorithm \code{cubBayesLattice\_g}:  a multivariate normal probability, the Keister's example, and an option pricing example.  Four different error tolerances, $\varepsilon$,  were set for each example, with the tolerances chosen depending on the difficulty of the problem.  The sequences $\{\vx_i\}_{i=1}^\infty$ were the randomly shifted lattice node sequences supplied by GAIL. The accuracy of the algorithm differs based on the shift. For each integral,  each tolerance, and each of our stopping criteria---empirical Bayes, full Bayes, and generalized cross-validation---our algorithm was run for $100$ different random shifts.  For each test, the execution times were plotted against $\abs{\mu - \hmu}/\varepsilon$.  We expect $\abs{\mu - \hmu}/\varepsilon$ to be no greater than one, but hope that it is not too much smaller than one, which would indicate a stopping criterion that is too conservative.

>>>>

We use the following parameter values in the simulation: 
\begin{align*}
d' = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol points \cite{DicEtal14a,DicPil10a}.  The results are for 400 randomly chosen $\varepsilon$ in the interval $[10^{-5}, 10^{-2}]$ as shown in \figref{fig:MVN_Metern_d2b2}. In each run, the nodes are randomly scrambled.  We  observe the algorithm meets the error criterion 95\% of the time.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer, 
As shown in \figref{fig:MVN_Metern_d2b2}, computation time increases rapidly with $n$. 
The  maximum likelihood estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC_\vtheta$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

@@

Three integrals were evaluated using the GAIL algorithm \code{cubBayesLattice\_g}:  a multivariate normal probability, the Keister's example, and an option pricing example.  
The sequences $\{\vx_i\}_{i=1}^\infty$ were the randomly shifted lattice node sequences supplied by GAIL. 
For each integral,  each of our stopping criteria---empirical Bayes, full Bayes, and generalized cross-validation---our algorithm was run for $400$ different randomly chosen error tolerance. The error tolerances were chosen randomly in an interval depending on the difficulty of the problem. In each run, the nodes were also randomly shifted. The accuracy of the algorithm differs based on the shift.
For each test, the execution times were plotted against $\abs{\mu - \hmu}/\varepsilon$.  We expect $\abs{\mu - \hmu}/\varepsilon$ to be no greater than one, but hope that it is not too much smaller than one, which would indicate a stopping criterion that is too conservative.
